\textsc{RILS}-\textsc{ROLS} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-11-22 paper_resources/doc/paper.tex
Update

Candidate perturbations are being improved by performing OLS coefficient fitting (procedure \texttt{FitOLS}). This means that coefficients in any of linear combinations of current solution are being set by applying  ordinary least square method, already described in Section~\ref{sec:ols}.   
After this step, perturbations are usually better suited to given sample data $D_{tr}'$. 
Further, these perturbations are sorted w.r.t. $R^2$ metric in the descending order (line~\ref{line:orderR2}).

@@

In order to speed-up the fitness evaluation, we employ \emph{expression caching}. This means that values attached to expression trees or subtrees are persisted in key-value structure, such that the key is tree (subtree) textual representation, while the value is the $|D_{tr}'|$-sized vector of corresponding expression values on the sample training dataset $D_{tr}'$. (Of course, once the $D_{tr}'$ changes, which happens relatively rarely, the whole cache is cleared.) 
Caching is performed in a partial way -- when determining the value of a given expression $T$, it is not required to find the exact \emph{hit} inside the cache. So, if some subexpression (subtree) of $T$ is present in the cache, its value will be reused and further combined to calculate the whole fitness function value.

For example, let us have an expression $T = y^2 \cdot (x+y)/z - sin(x+y)$, sample training dataset $D_{tr}'=\{([1, 2, 5], 7), ([3, 4, 3], 5), ([4, 5, 3], 6), ([6, 7, 4], 3), ([3, 3, 6], 2)\}$ (here, input feature vector is labeled by $[x, y, z]$), and expression cache consisting of the following key-value entries $cache=\{(x+y, [3, 7, 9, 13, 6]), (y^2, [4, 16, 25, 49, 9])\}$. 
Expression $T$ does not need to be fully evaluated since some of its parts are inside the cache: $y^2$ and $x+y$. (Note that variables alone do not need to enter the cache, since they are already available as columns of  $D_{tr}'$.)
Each newly evaluated (sub)expression (except constant or variable alone) enters the cache. In this example, the new entries will correspond to keys $(x+y)/z$, $y^2 \cdot (x+y)/z$, $sin(x+y)$ and $y^2 \cdot (x+y)/z - sin(x+y)$.  
The maximal number of cache entries is set to 5000. Once this number is reached, the cache is cleared.

@@

 \fxnote{Marko: stigao do ovog dijela...}

>>>>

Candidate perturbations are being improved by performing OLS coefficient fitting (procedure \texttt{FitOLS}). This means that coefficients in any of linear combinations of current solution are being set by applying  ordinary least square method, already described in Section~\ref{sec:ols}.    \fxnote{Marko: Aco, dodaj ono oko  nelinearnog dijela i pordešavanja koeficijenata koji se traže u predefinisanom skupu vrijednosti...}
After this step, perturbations are usually better suited to given sample data $D_{tr}'$. 
Further, these perturbations are sorted w.r.t. $R^2$ metric in the descending order (line~\ref{line:orderR2}).

@@

In order to speed-up the fitness evaluation, we employ \emph{expression caching}. This means that values attached to expression trees or subtrees are persisted in key-value structure, such that the key is tree (subtree) textual representation, while the value is the $|D_{tr}'|$-size  vector of corresponding expression values on the sample training dataset $D_{tr}'$. (Of course, once the $D_{tr}'$ changes, which happens not that frequently, the whole cache is cleared.) 
Caching is performed in a partial way -- when determining the value of a given expression $T$, it is not required to find the exact \emph{hit} inside the cache. So, if some subexpression (subtree) of $T$ is present in the cache, its value will be reused and further combined to calculate the whole fitness function value.

For example, let  $T = y^2 \cdot (x+y)/z - \sin(x+y)$ be an expression, and $D_{tr}'=\{([1, 2, 5], 7), ([3, 4, 3], 5), ([4, 5, 3], 6), ([6, 7, 4], 3), ([3, 3, 6], 2)\}$ be a sample training dataset (here, input feature vector is labeled by $[x, y, z]$); let expression cache consisting of the following key-value entries $cache=\{(x+y, [3, 7, 9, 13, 6]), (y^2, [4, 16, 25, 49, 9])\}$.  
Expression $T$ does not need to be fully evaluated since some of its parts are inside the cache: $y^2$ and $x+y$. (Note that single variables do not need to enter the cache, since they are already available as columns of  $D_{tr}'$.)
Each newly evaluated (sub)expression (except constant or variable alone) enters the cache. In this example, the new entries will correspond to keys $(x+y)/z$, $y^2 \cdot (x+y)/z$, $\sin(x+y)$ and $y^2 \cdot (x+y)/z - \sin(x+y)$.  
The maximal number of cache entries is set to 5000. Once this number is reached, the cache is cleared.

@@
